\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[numbered]{bookmark}
\usepackage[colorinlistoftodos]{todonotes}

\title{SOA Analysis and Simulations}

\author{JKP}

\date{\today}

\begin{document}
\maketitle

\input{common-part}

\section{ASE -- Gaussian Approximation}

\textbf{Reference: EE348 notes and Agrawal}

\textbf{One-sided ASE noise psd per polarization}
\begin{equation}
S_{sp}(\nu) = (G-1)n_{sp}h\nu
\end{equation}

Noise figure
\begin{equation}
F_n = 2n_{sp}\frac{G-1}{G} \approx 2n_{sp}
\end{equation}
The input coupling $f$ reduces the noise figure i.e., $F_n' = F_n/f$. This is a problem in SOAs and thus they present higher noise figure.

Signal is in one polarization and noise may appear in two polarization with same power.
\begin{align} \nonumber
I(t) &= R|\sqrt{G}E_s(t) + E_n(t)|^2 + |E_n(t)|^2 \\
& = R\Big(G|E_s(t)|^2 + \sqrt{G}E_s(t)E_n^*(t) + \sqrt{G}E_s^*(t)E_n(t) + 2|E_n(t)|^2\Big)
\end{align}
If a polarizer is used, then the noise component on the orthogonal polarization is avoided. More generally we can write
\begin{equation}
I(t) = R\Big(G|E_s(t)|^2 + \sqrt{G}E_s(t)E_n^*(t) + \sqrt{G}E_s^*(t)E_n(t) + N_{pol}|E_n(t)|^2\Big)
\end{equation}
where $N_{pol}$ is the number of noise polarizations. Number of noise polarizations i.e., use of polarizer, doesn't affect signal-spontaneous beat noise.

\subsection{Signal-spontaneous beat noise}
Signal-spontaneous beat noise has a bilinear term of form $\sqrt{G}E_s(t)E_n(t)$. Assuming signal is real and CW over a certain interval, which is consistent with a PAM signal:

\begin{align} \nonumber
I(t) = RG|E_s(t)|^2 + 2\sqrt{G}RE_s(t)\mathrm{Re}\{E_n(t)\} + RP|E_n(t)|^2
\end{align}

Therefore, the signal-spontaneous beat noise has variance
\begin{equation}
\sigma^2_{sig-sp, k} = 4R^2G\bar{P}_{s,k}S_{sp}\Delta f = 2R^2G\bar{P}_sN_0\Delta f
\end{equation}
where $\bar{P}_{s,k}$ is the average signal power at the $k$th level, and $S_{sp} = 2N_0$, where $N_0$ is the equivalent one-sided baseband ASE PSD.

The signal-spontaneous beat noise, which is dominant in well-designed receivers, is not affected by the number of noise polarizations i.e., whether a polarizer is used. Furthermore, it's independent of the band-pass optical filter bandwidth.

\subsection{Spontaneous-spontaneous beat noise}

Spontaneous-spontaneous beat noise PSD is not white; it decays linearly with frequency. At DC the PSD is $2N_{pol}R^2S_{sp}^2\Delta\nu_{opt}$. However, assuming $\Delta f << \Delta\nu_{opt}/2$ we can write

\begin{equation}
\sigma^2_{sp-sp} = 2N_{pol}R^2S_{sp}^2\Delta\nu_{opt}\Delta f = \frac{1}{2}N_{pol}R^2N_0^2\Delta\nu_{opt}\Delta f 
\end{equation}

\subsection{Power Penalty vs Noise Figure}

The signal-spontaneous beat noise is dominant, and the BER is determined by the argument of the Q-function
\begin{equation}
\frac{\Delta P}{\sigma_{sig-sp}} \propto \frac{GP}{\sqrt{G^2PF_n}} = \sqrt{\frac{P}{F_n}}
\end{equation}
Thus, 1-dB increase in the noise figure corresponds in 1-dB power penalty.

\section{ASE -- Karhunen-Loéve Series Expansion for MGF Calculation} 

\subsection{Definitions}
\begin{itemize}
	\item $x(t)$ signal component after amplifier;
	\item $w(t)$ ASE noise;
	\item $N_0$ baseband equivalent ASE noise PSD i.e., $N_0 = 2S_{sp}$.
	\item $e(t) = x(t) + w(t)$ electrical field after amplifier;
	\item $h_o(t) \leftrightarrow H_o(f)$ optical bandpass filter;
	\item $h_e(t) \leftrightarrow H_e(f)$ electrical baseband filter;
	\item $e_o(t) = (x(t) + w(t))\ast h_o(t) = s(t) + n(t)$ electric field after  optical bandpass filter;
	\item $y(t) = |e_o(t)|^2\ast h_e(t)$ electric signal after filtering (doesn't include thermal noise);
\end{itemize}

\subsection{Karhunen-Loéve Series Expansion in the Frequency Domain -- Less accurate and requires more eigenvalues}
Electric signal after filtering (doesn't include thermal noise) as a function of the Frequency-domain filter responses
\begin{align} \nonumber
y(t) &= |e(t)\ast h_o(t)|^2\ast h_e(t) \\ \nonumber
& = \mathcal{F}^{-1}\{[E(f)H_o(f)\ast E(-f)^*H_o(-f)^*]H_e(f)\} \\
& = \iint_{-\infty}^{\infty} E(f_1)E^*(f_2)K(f_1, f_2)e^{-j2\pi(f_1-f_2)t}df_1df_2
\end{align}
where $K(f_1, f_2) = H_o(f_1)H_e(f_1-f_2)H_o^*(f_2)$ is obtained simply by change of variables.

The eigenfunctions of the Fredholm integral equation form a complete set of orthonormal basis functions, and the corresponding eigenvalues $\lambda_n$ are real and ordered as $\lambda_1 \geq \lambda_2 \geq \ldots$.
\begin{equation}
\int_{-\infty}^{\infty} K(f_1, f_2)\phi(f_1)df_1 = \lambda\phi(f_2)
\end{equation}

Given the completeness of the eigenfunctions we can use the expansion:
\begin{equation}
E(f)e^{j2\pi ft} = \sum_{n=1}^{\infty}e_n(t)\phi_n(f)
\end{equation}
where 
\begin{equation}
e_n(t) = x_n(t) + w_n(t) = \int_{-\infty}^{\infty}E(f)\phi^*_n(f)e^{j2\pi ft}df
\end{equation}

Using this expansion to solve for $y(t)$ results in
\begin{equation}
y(t) = \sum_{n=1}^{\infty}\lambda_n|x_n(t) + w_n(t)|^2
\end{equation}
The coefficients $w_n(t)$ are statistically independent and circularly symmetric complex random variables with zero mean and variance equal to the ASE spectral density per polarization mode $N_0$.
Therefore, $y(t)$ is a weighted sum of non-central $\chi^2$ independent random variables.

\subsubsection{Evaluating $\lambda_n$ and $x_n$}
Performing the integration in a band $[-F, F]$ where the optical filter $H_o(f)$ is significant results
\begin{equation}
\int_{-\infty}^{\infty} K(f_1, f_2)\phi(f_1)df_1 \approx \int_{-F}^{F} K(f_1, f_2)\phi(f_1)df_1 \approx \lambda\phi(f_2)
\end{equation}
This can be approximated by the \textbf{Gauss-Legendre quadrature rule}
\begin{equation}
\sum_{m=1}^{M_e}w_mK(\nu_m, f_2)\phi_n(\nu_m) \approx \lambda_n\phi_n(f_2)
\end{equation}
where the weights $w_m$ are always positive.

Making $f_2 = \nu_n, n = 1,\ldots, M_e$
\begin{equation}
\sum_{m=1}^{M_e}w_mK(\nu_m, \nu_n)\phi_n(\nu_m) \approx \lambda_n\phi_n(\nu_n), n = 1,\ldots, M_e
\end{equation}
In matrix notation:
\begin{equation} \label{eq:matrix-form}
\mathcal{K}\mathcal{W}\Phi = \Phi\Lambda
\end{equation}
where 
\begin{align}
\mathcal{K}_{ij} = K(\nu_i, \nu_j) \\
\mathcal{W}_{ij} = w_i\delta_{ij} \\
\Phi_{ij} = \phi_j(\nu_i) \\
\Lambda_{ij} = \lambda_i\delta_{ij}
\end{align}
The columns of $\Phi$ are the eigenvectors of $\mathcal{K}\mathcal{W}$. But $\mathcal{K}\mathcal{W}$ is not Hermitian, so instead we can write $A = \mathcal{W}^{1/2}\mathcal{K}\mathcal{W}^{1/2}$, which is Hermitian. $\mathcal{W}^{1/2} = \mathrm{diag}\{\sqrt{w_1},\ldots,\sqrt{w_{M_e}}\}$. So instead of solving the eigenvector problem for \eqref{eq:matrix-form} it's more convenient to solve it for

\begin{equation} \label{eq:matrix-form}
\mathcal{W}^{1/2}\mathcal{K}\mathcal{W}\Phi = \mathcal{W}^{1/2}\Phi\Lambda = AB = B\Lambda
\end{equation}
where $A = \mathcal{W}^{1/2}\mathcal{K}\mathcal{W}^{1/2}$, and $B = \mathcal{W}^{1/2}\Phi$.

Once the eigenvectors $B$ are determine we can solve for $\Phi = \mathcal{W}^{-1/2}B$.

Knowing the sampled $\phi_n(\nu_m)$, we can interpolate it, and finally calculate $x_n$ by:
\begin{equation}
x_n(t) = \int_{-\infty}^{\infty}X(f)\phi_n^*(f)e^{j2\pi ft}df
\end{equation}

\subsection{Karhunen-Loéve using Fourier Series Basis}

Assume $x(t)$ is periodic with period $NT$, where $N$ is the number of symbols and $T$ is the symbol period.

Both signal and noise are expanded using the Fourier Series Basis functions (complex exponentials). 
\begin{align} \label{fourier-series}
& x(t) = \sum_{m=-M}^M x_me^{j2\pi n t/(NT)} \\
& w(t) = \sum_{m=-M}^M w_me^{j2\pi n t/(NT)} \\
\end{align}
where $L$ is large enough to account for great part of the energy in $x(t)$. This method sacrifices efficiency for simplicity. Other methods can expand the noise and signal in different bases and thus reduce the dimensionality required \cite{forestieri}.

\begin{align} \nonumber
y(t_k) = (x_k + n)^HK(x_k + n)
\end{align}
where
\begin{equation}
K_{ij} = K\bigg(\frac{i-M-1}{NT}, \frac{j-M-1}{NT}\bigg)  = K(f_i, f_j) = H_o(f_i)H_e(f_i-f_j)H_o^*(f_i);
\end{equation}
$(\cdot)^H$ denotes Hermitian transpose (conjugate transpose); $n$ is a column vector of zero-mean Gaussian random variables with variance $N_0/N$, since it corresponds to the Fourier Series coefficients of the periodic extension of the ASE noise; and $x_k$ is a column vector given by

\begin{equation}
x_{i, k} = x_{i - M - 1}e^{j2\pi(i-M-1)t_k/(NT)}
\end{equation}
where $x_{i - M - 1}$ are the Fourier series coefficients of the periodic extension of period $NT$ of $x(t)$ \eqref{fourier-series}.

Let 
\begin{equation}
c_k = U^Hx_k
\end{equation}
where $K = U\Lambda U^H$. $K$ is Hermitian and thus it only has real eigenvalues.

Therefore,
\begin{equation}
y(t_k) = (c_k + z)^H\Lambda (c_k + z) = \sum_{i = 1}^{M_e} \lambda_i|c_{i, k}^2 + z_i|^2
\end{equation}

And the MGF is given by
\begin{equation} \label{ASE-MGF}
\Psi_y(s) = \prod_{i =1}^{M_e}\frac{1}{(1-\lambda_i\sigma^2s)^{N_{pol}}}\exp\bigg(\frac{\lambda_i|c_{i,k}|^2s}{1-\lambda_i\sigma^2s}\bigg)
\end{equation}
where $\sigma^2 = N_0/{NT}$

We can calculate the first two moments of $y$ using \eqref{ASE-MGF}
\begin{align}
& \mu_y = \frac{d}{ds}\Psi_y(s)\bigg|_{s=0} = \Psi_y(s)\frac{d}{ds}\log\Psi_y(s)\bigg|_{s=0} = \sum_{i=1}^{M_e} \lambda_i(N_{pol}\sigma^2 + |x_n|^2) \\
& \sigma_{y}^2 = \frac{d^2}{ds^2}\Psi_y(s)\bigg|_{s=0} -\mu_y^2 = \Psi_y(s)\frac{d^2}{ds^2}\log\Psi_y(s)\bigg|_{s=0} = \sum_{i =1}^{M_e} N_{pol}(\lambda_i\sigma^2)^2 + 2\sigma^2\lambda_i^2|c_{i,k}|^2
\end{align}
the derivatives are calculated with respect to the derivatives of $\log\Psi_y(s)$, which is easier to obtain, as calculated in \eqref{mgf-pdf-derivatives}. Thus the conditional probabilities of $y$ can be approximated as a Gaussian distribution such that $p(y|a_i) \sim \mathcal{N}(\mu_y, \sigma_y^2)$.

\subsection{pdf and tail probabilities including thermal noise and using the saddlepoint approximation}

Including thermal noise in \eqref{ASE-MGF}
\begin{equation}
\Psi_y(s) = \bigg[\prod_{i =1}^{M_e}\frac{1}{(1-\lambda_i\sigma^2s)^{N_{pol}}}\exp\bigg(\frac{\lambda_i|c_{i,k}|^2s}{1-\lambda_i\sigma^2s}\bigg)\bigg]e^{\frac{1}{2}\sigma_{T}^2s^2}
\end{equation}

\subsubsection{pdf}
From the inversion equation we can write
\begin{align} \label{mgf-pdf-derivatives}
& K(s,x) = \log (\Psi(s)) - sx = \sum_{i = 1}^{M_e}\bigg( -N_{pol}\log(1-\lambda_i\sigma^2s) + \frac{\lambda_i|c_{i,k}|^2s}{1-\lambda_i\sigma^2s}\bigg) +\frac{1}{2}\sigma_T^2s^2 - sx  \\ \nonumber
&\frac{\partial K(s,x)}{\partial s} = \sum_{i = 1}^{M_e}\bigg( N_{pol}\frac{\lambda_i\sigma^2}{1-\lambda_i\sigma^2s} + \frac{\lambda_i|c_{i,k}|^2s}{(1-\lambda_i\sigma^2s)^2}\bigg) + \sigma_T^2s - x \\ \nonumber
& = \sum_{i = 1}^{M_e}\bigg( \frac{N_{pol}\lambda_i\sigma^2(1-\lambda_i\sigma^2s) + \lambda_i^2|c_{i,k}|^2s}{(1-\lambda_i\sigma^2s)^2}\bigg) + \sigma_T^2s - x \\
& = \sum_{i = 1}^{M_e}\bigg( \lambda_i\frac{N_{pol}\sigma^2(1-\lambda_i\sigma^2s) +|c_{i,k}|^2s}{(1-\lambda_i\sigma^2s)^2}\bigg) + \sigma_T^2s - x \\
&\frac{\partial^2 K(s,x)}{\partial s^2} = \sum_{i = 1}^{M_e}\bigg( \frac{N_{pol}(\lambda_i\sigma^2)^2}{(1-\lambda_i\sigma^2s)^2} + \frac{2\sigma^2\lambda_i^2|c_{i,k}|^2}{(1-\lambda_i\sigma^2s)^3}\bigg) + \sigma_T^2
\end{align}

\subsubsection{tail}
From the inversion equation we can write 
\begin{align}
& K(s,x) = \log \Big(\frac{\Psi(s)}{s}\Big) - sx =  \sum_{i = 1}^{M_e}\bigg( -N_{pol}\log(1-\lambda_i\sigma^2s) + \frac{\lambda_i|c_{i,k}|^2s}{1-\lambda_i\sigma^2s}\bigg) -\log |s| +\frac{1}{2}\sigma_T^2s^2 - sx  \\
&\frac{\partial K(s,x)}{\partial s} = \sum_{i = 1}^{M_e}\bigg( \lambda_i\frac{N_{pol}\sigma^2(1-\lambda_i\sigma^2s) +|c_{i,k}|^2s}{(1-\lambda_i\sigma^2s)^2}\bigg) - \frac{1}{s} + \sigma_T^2s - x \\
&\frac{\partial^2 K(s,x)}{\partial s^2} = \sum_{i = 1}^{M_e}\bigg( \frac{N_{pol}(\lambda_i\sigma^2)^2}{(1-\lambda_i\sigma^2s)^2} + \frac{2\sigma^2\lambda_i^2|c_{i,k}|^2}{(1-\lambda_i\sigma^2s)^3}\bigg) + \frac{1}{s^2} + \sigma_T^2
\end{align}
\textbf{Note that $K(s, x)$ is different from the pdf calculation.}


\section{Crosstalk in SOA}

Crosstalk occurs because the carrier lifetime of SOAs is in the order of the symbol rate. Thus, the SOA reacts fast enough to make fluctuations in the input power cause fluctuations in the gain.

Channel crosstalk in the amplifier has two main origins. One is the carrier density modulation at beat frequencies of the incident signals. This crosstalk mechanism is strongly dependent on channel frequency separation, and can thus be avoided for relatively large channel intervals, e.g., greater than 10 GHz. The other factor is gain reduction induced by the other channels due to gain saturation \cite{soa-crosstalk-and-penalty}.

Consider a traveling wave amplifier with negligible residual facet reflectivities. \textbf{When the input data rate is much smaller than the reciprocal of the lifetime}, the carrier density attains steady state within a small fraction of the bit period. For this case, the rate equation reduces to a \textbf{steady-state charge conservation equation}, and can be solved to yield the amplifier gain $G$ as a function of the input power $P_{in}$ \cite{soa-crosstalk}:
\begin{equation} \label{gain-saturation}
P_{in} = \frac{P_{sat}}{G-1}\ln\frac{G_0}{G}
\end{equation}
where $P_{sat}$ is the saturation power, and $G_0$ is the unsaturated gain

Here we will assume that \eqref{gain-saturation} holds even when the input data rate is higher than the reciprocal of the carrier lifetime. The difference is that the input power will be considered the average of the $N_s = B\tau_c$ symbols over $M_{ch}$ channels was done in \cite{soa-crosstalk-inc-lifetime}.

Thus, the first moments of the input power are given by
\begin{align}
& E(P_{in}) = \frac{1}{N_s}\sum_{i = 1}^{N_sM_{ch}} E(P_i) = M_{ch}E(P_i) \\
& E(|P_{in}|^2) = \frac{1}{N^2_s}\sum_{i, j = 1}^{N_sM_{ch}} E(P_iP_j) = \frac{M_{ch}}{N_s}\Big(E(P_i^2) + E(P_i)^2\Big) + M_{ch}^2E(P_i)^2 \\
& \text{Var}(P_{in}) = \frac{M_{ch}}{N_s}\Big(E(P_i^2) + E(P_i)^2\Big)
\end{align}
where $P_i$ is the power of the $i$th symbol. If $N_sM_{ch}$ is high $P_{in}$ can be modeled as a Gaussian random variable. Note that the variance reduces with $N_s$, but increases with $M_{ch}$. Thus, only increasing the carrier lifetime helps reducing the crosstalk penalty.


\subsection{Probability Mass Function of $P_{in}$}
Since $P_in$ comprises of a sum of $N_sM_{ch}$ symbols with power $P_i$, $P_{in}$ is a discrete random variable.

For binary signals we can model $P_{in}$ simply as a binomial distribution:
\begin{equation}
p\Big(P_{in} = k_1P_1 + (N_sM_{ch} - k_1)P_0\Big) = \frac{1}{2^{N_sM_{ch}}}\binom{N_sM_{ch}}{k_1}
\end{equation}
which basically calculates the probability of having $k_1$ '1s' (and consequently $N_sM_{ch}-k_1$ '0s') in $N_sM_{ch}$ Bernolli trials.

For a M-PAM signal we could write:
\begin{equation}
P_{in} = n_0P_0 + n_1P_1 + \ldots + n_{M-1}P_{M-1}
\end{equation}
where $n_i$ is the number of times the $i$th symbol was ``drawn'' in $N_sM_{ch}$ ``trials''. The number of different levels combinations is given by\footnote{This is equivalent to solve the problem: how many non-negative solutions are there in $\sum_{i = 1}^r n_r = N$? The number of non-negative solutions is given by $\binom{N+r-1}{r-1}$. The number of positive solutions is given by $\binom{N-1}{r-1}$. This problem is solved by imagining all possible ways to separate $N$ objects with $r$ containers.}
\begin{equation}
\#_{solutions} = \binom{N_sM_{ch}+M-1}{M-1}
\end{equation}


\begin{thebibliography}{9}
	
	\bibitem{ber-saddlepoint-approx} Carl Helstrom; ``Performance Analysis of Optical Receivers by the Saddlepoint Approximation,'' \emph{IEEE Transactions on Communications}, 1979.
		
	\bibitem{agilent-RIN-measurement} Product Note 86100-7; ``Digital Communication Analyzer (DCA), Measure Relative Intensity Noise (RIN),'' Agilent Technologies. 
	
	\bibitem{forestieri} E. Forestieri, M. Secondini; ``On the error probability evaluation in lightwave systems with optical amplificaiton,'' \emph{Journal of Lightwave Technol.}, 2009.
	
	\bibitem{soa-crosstalk-and-penalty} Kyo Inoue; ``Crosstalk and Its Power Penalty in Multichannel Transmission due to Gain Saturation in a Semiconductor Laser Amplifier,''	1989.
	
	\bibitem{soa-crosstalk} R. Ramaswami and P. Humblet; ``Amplifier induced crosstalk in multichannel optical networks,'', 1990.
	
	\bibitem{soa-crosstalk-inc-lifetime} S. Xu, J. Khurgin, I. Vurgaftman, J. Meyer; ``Reducing crosstalk and signal distortion in wavelength-division multiplexing by increasing carrier lifetimes in semiconductor optical amplifiers,'' 2003.
\end{thebibliography}

\end{document}